Impact of switching bug trackers: a case study on a medium-sized open source project

### 2019-07-26 article/article.tex
Final version of the ICSME paper.

%\documentclass[runningheads]{llncs}
\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
%
\usepackage{graphicx}
% If possible, figure files should be included in EPS format.
%
\usepackage{amsmath}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref,xcolor}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Impact of switching bug trackers: a case study on a medium-sized open source project}
%
%\author{Th\'eo Zimmermann\\
%$\Pi R^2$, Inria and
%IRIF, Paris-Diderot University\\
%Paris, France\\
%Email: theo@irif.fr
%\and
%Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
%Paris School of Economics\\
%Paris, France}
%
\maketitle
%
\begin{abstract}
For most software projects, the bug tracker is an essential tool. In open source development, this tool plays an even more central role as it is generally open to all users, who are encouraged to test the software and report bugs. Previous studies have highlighted the act of reporting a bug as a first step leading a user to become an active contributor.

@@

We first report on the switch itself, including the migration of preexisting bug reports. Then we analyze data from before and after the switch using a Regression on Discontinuity analysis, a novel methodology in the context of empirical software engineering. We complete this quantitative analysis with qualitative data from interviews with developers.
We show that the switch induces an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.
\end{abstract}

@@

\begin{IEEEkeywords}
bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, Regression on Discontinuity, interviews
\end{IEEEkeywords}
%
%
%
\section{Introduction}

Bug reporting is an essential part of software development.
In the context of an open source project, the bug reporting and fixing process is generally done on a public bug tracking platform, and, in the absence of paid testers, it depends a lot on the good will of independent users. Therefore, it involves a strong social component.

However, having an open bug tracking system is not enough to get the bugs reported. In addition to the software having enough users, the process of bug reporting must be easy and appealing. Therefore, creating an environment that eases bug reporting and makes it more appealing may lead to an increase in bug reporting activity (from both developers and independent users), which is well associated with an increase in software quality. First, if the software has not changed, receiving more bug reports indicates that more defects have been found, and not a sudden apparition of defects (Linus's Law ``given enough eyeballs, all bugs are shallow''~\cite{raymond1999cathedral,van2009shallow}). Moreover, assuming that developers are equipped to cope with the increased number of incoming reports~\cite{anvik2005coping, davidson2011coping}, even duplicate reports are not necessarily harmful~\cite{Bettenburg2008}. Second, more activity on the bug tracker can also mean users are helping to reproduce bugs, produce traces, etc. and thus are working with the developers to get the bugs fixed~\cite{breu2010information}. More generally, reporting new bugs and discussing existing bug reports has been shown to be an important step on the path to becoming an active contributor of an open source project~\cite{jensen2007role, nakakoji2002evolution, ye2003toward}.

Despite the importance of the bug tracking environment that we have just highlighted, the impact of a change in this environment has rarely been studied. It is generally difficult to compare the number of bug reports in different environments because it is likely to depend  not just on the reporting environment but also on each particular software and its development dynamics, which would make such an analysis not compelling. To do a valid comparison, one would ideally need to compare two bug reporting environments for the same software, used by the same users, to report the same bugs, in the same period. This is of course impossible. A context approaching this ideal scenario would be to compare the number of bug reports of a particular software before and after the switch to a new environment. Yet, substantial changes in the environment of bug reporting (e.g. changes of bug tracker platform) are rare because once a bug tracking platform is in place, any change comes with a significant cost.

In this paper, we study the case of the Coq proof assistant, a medium-sized open source software project. We use the switch, from Bugzilla to GitHub, of the project's bug tracker, to analyze how a change in the bug reporting environment affects the number of bugs, reporters, bug comments and commentators. Using a temporal Regression on Discontinuity design (RDD)~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression,hausman2018regression}, we find strong evidence that the change of platform increased the number of bug reports and comments by developers, and increased the number of distinct commentators each week among non-developers. Our results suggest that switching from Bugzilla to GitHub made the bug reporting process easier and more appealing but also opened the software development to more external contributors and stimulated discussion around bugs (notably between developers and non-developers).

The contributions of this paper are practical, empirical, and methodological. First, we improved an existing bug tracker migration tool to handle thousand of bug reports while preserving meta-data, and we helped the Coq development team migrate their bug tracker from Bugzilla to GitHub. Second, we analyze the causal impact of this switch on the bug tracking activity. Third, we introduce to the field of software engineering a state-of-the-art method for quantitative public policy evaluation: Regression in Discontinuity Design. In particular, we explain this method to give the basic intuition behind; we point the reader to the both introductory and state-of-the-art literature and we illustrate one application.

The rest of this paper is organized as follows: In Sec.~\ref{related-work}, we discuss related work. In Sec.~\ref{context}, we present the Coq project and the context of the switch. In Sec.~\ref{research_questions}, we present our research questions. In Sec.~\ref{migration}, we explain the migration process. In Sec.~\ref{data}, we describe the data collection and pre-processing method, and we list the variables of interest. In Sec.~\ref{descriptive-stats}, we give a preliminary view of the bug tracking activity. In Sec.~\ref{methodology}, we explain the Regression on Discontinuity methodology, of which we present the results in Sec.~\ref{results}. We interpret the results and relate them to our hypotheses in Sec.~\ref{discussion}. We discuss the threats to validity and  present robustness checks in Sec.~\ref{threats}. We conclude in Sec.~\ref{conclusion}.

@@

While there is a very large literature on many aspects of bug reporting (see~\cite{strate2013literature, zhang2016literature} for literature reviews), there is a lack of literature measuring the influence of the bug reporting environment on the bug reporting activity or on any other aspects of software development. To our knowledge, there is no previous work measuring the impact of a change in bug reporting environment. More generally, there is little literature that compare bug trackers.

@@

A simple, but quite limited, way to compare bug trackers is to look at the features that each of them proposes.

Karre et al.~\cite{karre2017does} compared 31 bug trackers feature-wise, and gathered these tools in four clusters. Bugzilla belongs to the cluster of bug trackers with many features (together with RedMine and Mantis), while GitHub belongs to a cluster of bug trackers attached to source code management systems (together with Savannah and BitBucket).
%Bugzilla has 20 of the 23 features they listed, and scores the highest number of features (ex. aeqo with RedMine). They are closely followed by Mantis, and two other bug trackers, with 19 features. GitHub only has 15 features out of the 23 listed (overall 11 listed systems have more features).

In the case of the Coq project, the loss of some specific features from Bugzilla was one of the arguments against the switch. Nonetheless, GitHub has some features that Bugzilla hasn't, and it is clearly not the number of features that determines which software is better or more adapted to the specific use case.

%We will note that GitHub has very recently added a number of new features to the issue tracker, but these were not present at the time of the switch (just-in-time duplicate retrieval, issue deletion, transfer and pinning, support for multiple issue templates, etc.).

Additionally, Karre et al. recommended some missing features, including better support for migrating bug tracking data from a system to another. Our work has contributed to improve the situation in that sense. Nonetheless, it would be preferable if there existed a standard structured format for storing bug tracking information and all major bug tracking systems supported exporting to and importing from this format.

Abaee and Guru~\cite{abaee2010enhancement} listed the features of four commercial bug trackers, and proposed their own bug tracker with some new features, but without any evaluation.

There is much more literature proposing new features to add to bug trackers, but they are rarely evaluating the impact of adding such features on the bug tracking activity, even when a prototype was presented.

Baysal et al.~\cite{baysal2013situational} identified the need for personalized issue tracking systems after interviewing twenty Mozilla developers. They proposed a Bugzilla extension to address this need, and gave a, mostly qualitative, assessment by interviewing developers using their tool in~\cite{baysal2014no}.

Bortis~\cite{bortis2016porchlight} developed a bug triaging tool and evaluated how users interacted with it. However, there was no evaluation of its impact in the context of an actual software project.

Just et al.~\cite{just2008towards} surveyed developers from three large open source projects (Apache, Eclipse, and Mozilla; all three projects use Bugzilla) to identify missing features and provided recommendation to design better bug tracking systems.
%They remark that bug trackers are often mere interfaces to bug databases and ask too much information from end-users without guiding them. They recommend providing guidance to bug reporters depending on their level of expertise. This issue is indeed something that we have noticed with Bugzilla and other bug trackers of the same category. Users are encouraged to self-triage and often do so inaccurately. On the contrary, GitHub only provides title and body fields to bug reporters and leaves triaging (labeling) to the contributors with write access. Nonetheless, the issue templates can be used to fill the bug report body field with advice on how to present a bug report. GitHub has recently added support for multiple issue templates which would allow to let the reporter choose between, for instance "My first bug report", "My first feature request", "Bug report (experienced reporter)", and "Feature request (experienced reporter)", with varied requirements and levels of explanations...

Our article goes beyond a descriptive or normative perspective and quantitatively measures the causal impact of the change of the bug tracking platform (a crucial part of the bug tracking environment) on various aspects of bug reporting activity.

@@

Another way of comparing bug tracking systems could have been to conduct large-scale studies of many software projects using various bug trackers and provide some clues of the differences induced by the use of the different bug tracking systems. However, there are no such comparative studies in the literature.

Sowe et al.~\cite{sowe2013multi} noted that most preceding literature had only been comparing few projects at a time, usually using a single bug tracking system. In their study, they addressed this in part by studying hundreds of projects, but they did not mention which bug trackers the projects they compared used.\looseness=-1

Bissyandé et al.~\cite{bissyande2013got} published the same year a larger scale investigation using ten thousand projects' bug tracking data, analyzing such things as the correlation between a project's success and its bug tracking activity. All of the projects studied were using the same bug tracker (GitHub's).

Francalanci and Merlo~\cite{francalanci2008empirical} analyzed the bug fixing process by studying closed bugs from nine open source projects (four using JIRA, four using SourceForge's bug tracker, and one using another bug tracker). They did not, however, try to correlate the bug tracking activity with the bug tracking system used.

%According to the survey in~\cite{karre2017does}, Bugzilla is the most used bug tracker (respondent were 32\% to use Bugzilla and 15\% to use GitHub). There is an obvious selection bias due to the choice of open source project contributors this survey was sent to. Overall, given the sheer number of projects hosted on GitHub, it does not make any doubt that the GitHub issue tracker is the most used bug tracker in number of projects. But this raw fact does not matter much, given that very small projects will use their bug tracker very little. 
\subsection{Switching developer tools}

While, to the best of our knowledge, there is no work evaluating the impact of switching bug trackers, some studies have been conducted to evaluate the impact of switching other developer tools. Since bug trackers are important developer tools, our article also relates and contributes to this literature.

De Alwis and Sillito~\cite{de2009software} surveyed developers to understand the perceived impact of moving from centralized to decentralized version control systems. As with bug trackers, the move is decided because of anticipated benefits: among them is the openness to external contributors. But the switch incurs challenges in the migration process (in particular, developers are very concerned to preserve the full history of the project) and some features are lost in the new tool (in particular, the monotonically increasing revision numbers).
Muşlu et al.~\cite{mucslu2014transition} performed a similar study based on interviews and a survey. They identified expectations and barriers and provided recommendations to developers wanting to conduct such a migration.

Squire~\cite{squire2015should} measured the impact of switching developer support channels from mailing lists and self-hosted forums to Stack Overflow, by first identifying the expectations of 20 projects which had officially decided on such a move, then by studying the impact on developer participation and response time for a selection of four such projects. On the other hand, Vasilescu et al.~\cite{vasilescu2014social} compared the behavior of the same users on the R mailing list and the Stack Overflow platform. Contrarily to a bug tracker which is expected to be centralized for a given project, support channels can be diverse. Project leaders can decide to abandon their mailing list to focus exclusively on Stack Overflow, but they do not decide to ``open'' a support channel on Stack Overflow: it can exist nevertheless and the switch can be organic.

@@

% XXXX justify more the medium-size qualification

The development is currently centered around the GitHub platform. The Coq repository counts more than 25,000 commits. Some of these commits date back from 1999, when the code received major architectural changes. The GitHub repository was initially created as a mirror of the official repository (which itself was migrated from SVN to git in 2013) and really started attracting the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015 (later renamed into the Coq Implementors Workshop and held every year in France). From then on, core developers started using pull requests for their own changes more and more often, and it became the norm around the beginning of 2017 when systematic continuous integration tests were introduced on every pull request. In the last two years, more than 2,200 pull requests have been opened.

@@

From 2007 to 2017, the Coq bug tracker platform was a self-hosted Bugzilla instance (before this, from 2001 to 2007 it was using JitterBug, a now discontinued bug tracking system, and before 2001 a simple mailing list). In 2017, in a context where all code changes were conducted through GitHub pull requests, the question of switching to GitHub's integrated bug tracker arose.
After some preliminary testing showing that migrating all preexisting issues was doable, the Coq development team decided to switch to GitHub issues on October 4th, 2017.

According to the developers we interviewed, the motivations for switching were:
\begin{itemize}
\item Consolidating the development tools on a single, integrated, platform: browsing between code, pull requests and bug reports is easier without having to switch websites. Furthermore, it means a common notification, mention, and assignment system.
\item GitHub's support for formatting and editing comments, cross-referencing and auto-closing issues, and more generally a more pleasant tool to use. On the other hand, many Coq developers perceived Bugzilla as unpleasant, old-fashioned, and very slow. It was a complex tool with many underused advanced features.
\item No need to administrate (maintain and update) the bug tracker. The previous system was perceived as failing often by several developers.
\item Easier to get started for newcomers (especially as many may already know GitHub and already have a GitHub account).
\end{itemize}

Again, according to the developers we interviewed, the principal risks associated with the switch of bug tracker were:
\begin{itemize}
\item Losing control of a critical tool to a private company and its future decisions. In particular, GitHub is not an open source platform (so cannot be cloned elsewhere) and backing up data from GitHub becomes even more important but not trivial to do.
\item Losing or corrupting bug tracking data during the transition from Bugzilla to GitHub. Even if, as we will see in Sec.~\ref{migration}, most of the data was correctly migrated, some meta-data had to be converted to text, and external links and documentation about the bug tracker could have become stale or inaccurate. %In practice, developers were quite happy about the result of the migration.
\item A few developers additionally mentioned risks of losing useful features from Bugzilla, or discouraging developers for whom it would be hard to adapt to the new tool.
\end{itemize}

\section{Research questions}
After a consideration of the motivations and keeping the possible risks in mind, we found particularly interesting and compelling to study the following research questions more in detail to give insights into the effect of a bug tracker switch
\label{research_questions}
%ambre: I think this section should be expanded a bit to add motivation for the hypotheses

%Given the advantages of the GitHub issue tracker that were listed above, one could expect the switch to:
%\begin{description}
%\item[H1] Increase the bug tracking activity, in terms of the number of reported bugs;
%\item[H2] Extend the pool of bug reporters and commentators;
%\item[H3] Increase the bug tracking activity, in terms of the amount of interaction between users and developers.
%\end{description}

\paragraph{RQ1: Impact on level of bug tracking activity.}
%Do not forget the "why"

\paragraph{RQ2: Impact on quality of bug tracking activity.}
\paragraph{RQ3: Impact on the audience of the bug tracker.}

All of this could lead in turn to a greater openness and transparency of the development and an increased engagement of users toward it, changing (and probably improving) the dynamics of a crucial part of software development.

@@

The only complicated part of the bug tracker switch was the migration of preexisting bug reports. We reused a tool by Andriy Berestovskyy\footnote{Tool available at \url{https://github.com/berestovskyy/bugzilla2github}.} which is designed to import Bugzilla reports (extracted as an XML dump) to GitHub using its REST API. The bugs are imported in an order designed to preserve numbers whenever possible. Bugs whose number is unavailable (e.g. because the number is already taken by a GitHub pull request) are postponed and renumbered. We implemented several improvements to make the tool better fit the needs of the Coq project:
\begin{itemize}
\item \textit{Allowing non-consecutive bug numbers:} The imported set of bug reports had some holes in the numbering due to deleted bugs. We use postponed bugs to fill the holes. This improvement has now been integrated upstream.
\item \textit{Saving a table of correspondence for renumbered bugs:} This was used later to redirect the old bug URLs to the new ones.
\item \textit{Using the GitHub issue import beta API and overcoming the GitHub rate limits:} Creating a new issue or a new comment through the normal GitHub REST API will trigger notifications (for people who are watching the repository or are mentioned in the issue thread). Therefore, GitHub chooses to impose a strict rate limit on these actions, which prevented using this tool for importing more than a few hundred bugs. Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one bug, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 bugs to just a few hours). Furthermore, using this API allows to keep the dates of imported bug reports and comments, which is very useful to this study. We didn't manage to import correct closing dates for every bug report, so we will not study the impact of the bug tracker switch on the time it takes to close a bug.

This improvement has now been integrated upstream as an optional setting.
\end{itemize}

This tool required a correspondence table between Bugzilla and GitHub accounts. Ours was created by manually matching 217 (out of 686) Bugzilla accounts to 175 GitHub accounts,\footnote{The correspondence table can be seen at \url{https://frama.link/vRV9etCd}.} the difference being due to the high prevalence of duplicate Bugzilla accounts. We gave priority to finding GitHub accounts for people who still had opened bug reports.
%The most efficient way to do the matching was by searching the name of the reporter + ``github'' on Google. We also sent e-mails to people we hadn't managed to match (and who still had opened bugs).

Once the switch was approved by the development team on October 4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of bug reports that would need to be renumbered. It was conducted %\footnote{More details on the process are available in a blog post at \url{http://theoz.im/bugzilla}. }
on October 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process). Only 502 out of 4900 bugs (whose numbers were below 1154) had to be renumbered. This number is quite low because these bugs are dating back from the JitterBug era and during the previous switch only the bugs that were still open had been migrated.
Due to the rate of pull request creation, if the switch had been delayed by a year, the number of renumbered bugs would have been closer to 3000.

@@

All the data for this study was extracted on January 17\textsuperscript{th}, 2019 using the GitHub GraphQL API \cite{github_graphql_API}. Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process). In the supplementary materials, we provide the extracted data as CSV files and a Jupyter notebook with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.

Migrated bugs and comments have their author information stored in the text, we extract it from there. Dates of bug and comment creation have been preserved which allows us to obtain them transparently for bugs before and after the migration. Bug reports and comments from the JitterBug era are missing information and were not fully migrated, therefore we do not consider any data from before 2008.

@@

\paragraph{Excluding specific reporters}
We have one specific reporter who is alone responsible for almost a quarter of all bug reports. To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and the comments they received from our analysis. Similarly, we also exclude the developer who was the main advocate of the bug tracker switch as it could have influenced his behavior before and after the switch.

\paragraph{Merging duplicate Bugzilla accounts}
We found that a significant number of users had created several accounts on the Bugzilla system. We merged them to avoid overestimating the rate of new reporters which is analyzed in the companion Jupyter notebook.

>>>>

\documentclass[conference]{IEEEtran}

\usepackage[noadjust]{cite}
\usepackage{balance}

\IEEEoverridecommandlockouts
%
\usepackage{graphicx}
% If possible, figure files should be included in EPS format.
%
\usepackage{amsmath}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref,xcolor}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Impact of switching bug trackers: a case study on a medium-sized open source project}
%
\author{Th\'eo Zimmermann (theo@irif.fr)\\
Universit\'e de Paris, IRIF, CNRS, F-75013 Paris, France\\
Inria, $\pi.r^2$ project-team
\and
Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
Paris School of Economics, F-75014 Paris, France}
%
\maketitle
%
\begin{abstract}
For most software projects, the bug tracker is an essential tool. In open source development, this tool plays an even more central role as it is generally open to all users, who are encouraged to test the software and report bugs. Previous studies have highlighted the act of reporting a bug as a first step leading a user to become an active contributor.

@@

We first report on the switch itself, including the migration of preexisting issues. Then we analyze data from before and after the switch using a regression discontinuity design, an econometric methodology imported from quantitative policy analysis. We complete this quantitative analysis with qualitative data from interviews with developers.
We show that the switch induces an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.
\end{abstract}

@@

\begin{IEEEkeywords}
bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, regression discontinuity design, RDD, interviews
\end{IEEEkeywords}
%
%
%
\section{Introduction}

Bug reporting is an essential part of software development.
In the context of an open source project, the bug reporting and fixing process is generally done on a public bug tracking platform, and, in the absence of paid testers, it depends a lot on the goodwill of independent users. Therefore, it involves a strong social component.

However, having an open bug tracking system is not enough to attract participants. In addition to the software having enough users, the process of opening an issue (be it bug report or feature request) must be easy and appealing. Therefore, creating a favorable bug tracking environment may lead to an increase in bug tracking activity (from both developers and independent users), which may result in an increase in software quality. First, the participation of more users is helpful to find a larger proportion of bugs~\cite{raymond1999cathedral,van2009shallow}. Second, assuming that developers are equipped to cope with the increased number of incoming reports~\cite{anvik2005coping, davidson2011coping}, even duplicate reports are not necessarily harmful~\cite{Bettenburg2008}. Third, more activity on the bug tracker can also mean users are helping to reproduce bugs, produce traces, etc. and thus are working with the developers to get the bugs fixed~\cite{breu2010information}. More generally, opening issues and discussing existing ones has been shown to be an important step on the path to becoming an active contributor of an open source project~\cite{jensen2007role, nakakoji2002evolution, ye2003toward}.

Despite the importance of the bug tracking environment that we have just highlighted, the impact of a change in this environment has rarely been studied, whether it is the bug tracking platform in full, a feature, or a policy.
Furthermore, since the choice of bug tracker is not independent from the characteristics of the project, a comparison across projects using different bug trackers would not be sufficient to draw causal conclusions.

In this paper, we study the case of the Coq proof assistant, a medium-sized open source software project.
The contributions of this paper are practical, empirical, and methodological. First, we improved an existing bug tracker migration tool to handle thousands of issues while preserving meta-data, and we helped the Coq development team switch their bug tracker from Bugzilla to GitHub. Second, we analyze the causal impact of this switch on the bug tracking activity through mining repository data, and we interpret the results through qualitative assessment based on interviews with developers. Third, we demonstrate the application of regression discontinuity design (RDD), an econometric method to derive causality, in the context of empirical software engineering; we explain the method to give the basic intuition behind it; we point the reader to introductory, basic, and state-of-the-art literature~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression,hausman2018regression}.

The rest of this paper is organized as follows: In Sec.~\ref{related-work}, we discuss related work. In Sec.~\ref{context}, we present the Coq project and the context of the switch. In Sec.~\ref{research_questions}, we present our research questions. In Sec.~\ref{migration}, we explain the migration process. In Sec.~\ref{data}, we describe the data collection and pre-processing, and we list the variables of interest. In Sec.~\ref{descriptive-stats}, we give a preliminary view of the bug tracking activity. In Sec.~\ref{methodology}, we explain our quantitative and qualitative methodology. We present quantitative results in Sec.~\ref{results}. We present qualitative results, interpret the quantitative results, and relate them to our research questions in Sec.~\ref{discussion}. We discuss threats to validity and present robustness checks in Sec.~\ref{threats}. We conclude in Sec.~\ref{conclusion}.

@@

While there is a very large literature on many aspects of bug reporting (see~\cite{strate2013literature, zhang2016literature} for literature reviews), there is a lack of literature measuring the influence of the bug reporting environment on the bug reporting activity, or on any other aspects of software development. To our knowledge, there is no previous work measuring the impact of a change in bug reporting environment. More generally, there is little literature that compares bug trackers.

@@

A simple, but quite limited, way to compare bug trackers is to look at the features that each of them proposes.
Karre et al.~\cite{karre2017does} compared 31 bug trackers feature-wise, and gathered these tools in four clusters. Bugzilla belongs to the cluster of bug trackers with many features (together with RedMine and Mantis), while GitHub belongs to a cluster of bug trackers attached to source code management systems (together with Savannah and BitBucket).
Abaee and Guru~\cite{abaee2010enhancement} listed the features of four commercial bug trackers, and proposed their own bug tracker with some new features, but with no evaluation.

Many papers propose new features to add to bug trackers, but they rarely evaluate the impact of adding such features, even when a prototype was presented.
Baysal et al.~\cite{baysal2013situational} identified the need for personalized issue tracking systems after interviewing twenty Mozilla developers. They proposed a Bugzilla extension to address this need, and gave a, mostly qualitative, assessment by interviewing developers using their tool in~\cite{baysal2014no}.
Bortis~\cite{bortis2016porchlight} developed a bug triaging tool and evaluated how users interacted with it. However, there was no evaluation of its impact in the context of an actual software project.
Just et al.~\cite{just2008towards} surveyed developers from three large open source projects (Apache, Eclipse, and Mozilla; all three projects use Bugzilla) to identify missing features and provided recommendation to design better bug tracking systems.

Our article goes beyond a descriptive or normative perspective and quantitatively measures the causal impact of the change of the bug tracking platform (a crucial part of the bug tracking environment) on various aspects of bug tracking activity, and gives qualitative insights for its interpretation.

@@

Another way of comparing bug tracking systems could have been to conduct large-scale studies of many software projects using various bug trackers, and derive some differences associated with the use of the various systems. However, there are no such comparative studies in the literature.

Sowe et al.~\cite{sowe2013multi} noted that most preceding literature had only been comparing few projects at a time, usually using a single bug tracking system. In their study, they addressed this in part by studying hundreds of projects, but they did not mention which bug trackers the projects they compared used.

Bissyand\'e et al.~\cite{bissyande2013got} published the same year a larger scale investigation using ten thousand projects' bug tracking data, analyzing such things as the correlation between a project's success and its bug tracking activity. All of the projects studied were using the same bug tracker (GitHub's).

Francalanci and Merlo~\cite{francalanci2008empirical} analyzed the bug fixing process by studying closed bugs from nine open source projects (four using JIRA, four using SourceForge's bug tracker, and one using another bug tracker). They did not, however, try to correlate the bug tracking activity with the bug tracking system.\looseness=-1

\subsection{Switching developer tools or processes}

While, to the best of our knowledge, no previous work has evaluated the impact of switching bug trackers, some studies have evaluated the impact of switching other developer tools or processes. Since bug trackers are important developer tools, our article also relates and contributes to this literature.

De Alwis and Sillito~\cite{de2009software} surveyed developers to understand the perceived impact of moving from centralized to decentralized version control systems. As with bug trackers, the move is decided because of anticipated benefits: among them is the openness to external contributors. But the switch incurs challenges in the migration process (in particular, developers are very concerned to preserve the full history of the project) and some features are lost in the new tool (in particular, the chronologically ordered revision numbers).
Muşlu et al.~\cite{mucslu2014transition} performed a similar study based on interviews and a survey. They identified expectations and barriers and provided recommendations to developers wanting to conduct such a migration.

Squire~\cite{squire2015should} measured the impact of switching developer support channels from mailing lists and self-hosted forums to Stack Overflow, and compared it with the expectations of projects having decided on such a move,
whereas Vasilescu et al.~\cite{vasilescu2014social} compared the behavior of the same users during the same period on the R mailing list and on Stack Overflow.

Alencar da Costa et al.~\cite{da2018impact} evaluated the impact of switching from traditional (long) release cycles to faster ones on the time it takes for a bug fix to get released. They did a quantitative assessment based on a case study on the Firefox project, and completed it by surveying developers.

\subsection{RDD in empirical software engineering}

While RDD is a well-established method in quantitative public policy evaluation, it has been seldom used in empirical software engineering.
The only two studies we know of \cite{zhao2017impact,trockman2018adding} analyze hundreds of thousands of GitHub repositories, with time as the rating variable.
In our paper, we show how this methodology can be used to derive causality from a smaller dataset (the impact of a change on a single active project).
We demonstrate precisely how to use this method, with specific details on the particular case of time series, how to visualize the results in a compelling way, what are the internal threats to validity, and how to mitigate them with appropriate robustness checks.

@@

The development is currently centered around the GitHub platform.
The GitHub repository, which counts more than 25,000 commits starting in 1999, was initially created as a mirror of the official SVN repository.
It attracted the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015.
Then, core developers began using pull requests for their own changes more frequently, and it became the norm around the beginning of 2017 when systematic continuous integration testing was introduced. Since 2017, more than 3,000 pull requests have been opened.

@@

From 2007 to 2017, the Coq bug tracker platform was a self-hosted Bugzilla instance (before this, from 2001 to 2007 it was using JitterBug, a now discontinued bug tracking system, and before 2001 a simple mailing list). In 2017, in a context where all code changes were conducted through GitHub pull requests, the question of switching to GitHub's integrated bug tracker arose.
After some preliminary testing showing that migrating all preexisting issues was doable, the Coq development team approved the switch to GitHub issues on Oct. 4\textsuperscript{th}, 2017.

According to the developers we interviewed, the motivations for switching were:
\begin{itemize}
\item Consolidating the development tools on a single, integrated, platform: browsing between code, pull requests and issues is easier without having to switch websites. Furthermore, it means a common notification, mention, and assignment system.
\item GitHub's support for formatting and editing comments, cross-referencing and auto-closing issues, and more generally a more pleasant tool to use. Many developers perceived Bugzilla as unpleasant, old-fashioned, and very slow. It was a complex tool with many underused advanced features.
\item No need to administrate (maintain and update) the bug tracker. Almost half of the developers complained about the previous bug tracker failing often.
\item Easier to get started for newcomers (especially as many may already know GitHub and have a GitHub account).
\end{itemize}

Again, according to the developers we interviewed, the principal risks associated with the switch of bug tracker were:
\begin{itemize}
\item Losing control of a critical tool to a private company and its future decisions. In particular, GitHub is not an open source platform (so cannot be cloned elsewhere) and backing up data from GitHub becomes even more important, but not trivial to do.
\item Losing or corrupting bug tracking data during the transition from Bugzilla to GitHub.
During the migration, some meta-data had to be converted to text, and external links and documentation about the bug tracker could have become stale or inaccurate.
\item A few developers additionally mentioned risks of losing useful features from Bugzilla, or discouraging some developers for whom it would be hard to adapt to the new tool.
\end{itemize}

\section{Research questions}

\label{research_questions}

Our research questions cover the spectrum of the possible impact such a switch could have on the bug tracking activity:

\paragraph{Impact on level of %bug tracking
activity (RQ1)}

Did the switch to the GitHub bug tracker, with the various expected benefits that such a switch would bring, impact the level of activity on the bug tracker? If so, who are the participants whose level of activity was impacted, and how can we explain such changes?

\paragraph{Impact on quality of %bug tracking
activity (RQ2)}

Did the switch to GitHub impact the quality of opened issues, and did it impact the quality of interactions between developers, and between developers and non-developers?

\paragraph{Impact on the audience of the bug tracker (RQ3)}

Did the switch to GitHub help onboard more new reporters and commentators? Did it increase the number of distinct non-developers that regularly take part to the bug tracking activity?

@@

As explained in Sec. \ref{context}, one of the main risks (and deterrents) to a switch of bug tracker was the possible loss or corruption of bug tracking and associated data. In particular, the Coq development team wanted preexisting issues to be migrated to the new bug tracker, while keeping the issue numbers as much as possible: indeed, these numbers are important because they are mentioned in many places (changelog, test-suite, other issues and pull requests, commits, external discussion boards, etc.).\looseness=-1

We reused and adapted a tool~\cite{bugzilla2github} which is designed to import Bugzilla reports (extracted as an XML dump) to GitHub using its REST API~\cite{github_REST_API}. The issues are imported in an order designed to preserve numbers whenever possible. Issues whose number is unavailable (e.g. because the number is already taken by a GitHub pull request) are postponed and renumbered. We implemented several changes to make the tool better fit the needs of the Coq project (the main improvements having now been integrated upstream):

\paragraph{Allowing non-consecutive numbers} The imported set of issues had some holes in the numbering due to deleted issues. We use postponed issues to fill the holes.
\paragraph{Saving a table of correspondence for renumbered issues} This was used later to redirect the old Bugzilla URLs.
\paragraph{Using GitHub's issue import API and overcoming GitHub's rate limits} Creating a new issue or a new comment through the normal GitHub REST API triggers notifications (for people who are watching the repository or are mentioned in the issue thread). Therefore, GitHub chooses to impose a strict rate limit on these actions, which prevented using this tool for importing more than a few hundred issues. Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one issue, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 issues to just a few hours). Furthermore, using this API allows to keep the dates of imported issues and comments, which is very useful to this study.

Once the switch was approved by the development team on Oct. 4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of issues that would need to be renumbered. It was conducted
on Oct. 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process). Only 502 out of 4900 issues (whose numbers were below 1154) had to be renumbered. This number is quite low because a lot of the most ancient issues had been deleted (during the previous bug tracker migration).
Due to the rate of pull request creation, if the switch had been delayed by a single year, the number of renumbered bugs would have been closer to 3000.

@@

All the data for this study was extracted on March 31\textsuperscript{st}, 2019 using the GitHub GraphQL API~\cite{github_graphql_API}. Using this API allows us to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process). In the supplementary materials, we provide the extracted data as CSV files and a Jupyter notebook with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps and the analyses, including the robustness checks presented in Sec.~\ref{threats}.

As was previously mentioned, the migration conserved the dates of issues and comments, which allows us to obtain them transparently for issues before and after the switch.
On the other hand, author information for migrated issues and comments had to be encoded in the text, and is thus extracted from there.
Finally, we do not consider any data from before 2008, because this data comes from a previous migration, and was not properly saved.

@@

\label{pre-processing}

\paragraph{Excluding specific reporters}
We have one specific reporter who is alone responsible for almost a quarter of all issues. To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, issues, and the comments they received from our analysis. Similarly, we also exclude the developer who was the main advocate of the bug tracker switch as it could have influenced his behavior before and after the switch.

\paragraph{Merging duplicate accounts}
Because of space constraints, we refer the reader to the supplementary Jupyter notebook for details on this step.

\paragraph{Removal of migration artifact comments}
The migration tool created artifact comments, which are easily identified because they are the comments that were posted at the exact same date and time as the corresponding issue. We exclude them from our analysis.
